Gametic Selection, Meiotic Drive, Sex Ratio Bias, and Transitions Between Sex Determination Systems

### 2017-03-02 LaTeX/Turnover.tex
position plots for w and y made
merging edits made from Sally's comments and new fig 3

Table \ref{tab:haplotype_growth} illustrates a number of key points about the invasion of neo-Y and neo-W mutations. 
For a neo-Y, invasion is driven by the fitness of male gametes and diploids, where the latter is weighted by the chance that a female egg will give rise to that diploid. 
For example, matings with $A$-bearing female gametes occur with probability $p_{X}^{\female} w_{A}^{\female}/\bar{w}_{H}^{\female}$. 
Since a neo-Y is always found in males, the allele frequencies at the neo-Y (\textbf{M}) locus only change in males. 
Therefore, invasion by a neo-Y allele does not involve any female diploid selection terms ($w_{g}^{\female}$). 
Similarly, invasion by a neo-W is driven by the fitness of female gametes and diploids and does not involve any direct selection in male diploids.
However, in the case of a neo-W, female diploids can result from matings with either an X-bearing or a Y-bearing sperm. 
In either case, the zygote will develop as a female due to the presence of a neo-W. 
For example, neo-W females will therefore inherit an $A$ from a male gamete with probability $\bar{p}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$, where $\bar{p}^{\male}=p_{Y}^{\male}q+p_{X}^{\male}(1-q)$. 
By contrast, females that do not carry the neo-W only result from matings with X-bearing sperm. 
They will therefore inherit an $A$ from a male gamete with probability probability $(1-q)\bar{p}_{X}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$. 
If the \textbf{A} locus is initially linked to the ancestral sex-determining locus, \textbf{X}, the frequency of the $A$ allele among X- and Y-bearing sperm can differ (equation \ref{eq:freq_diffs}). 
Thus, eggs with and without a neo-W differ in the frequency of $A$ alleles they obtain from mating with male gametes.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Heterogametic transitions with/without sex ratio change
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{Combination_Turnover}\\
\caption{
Heterogametic transitions from XY to ZW sex determination (neo-W frequency shown by black lines, panels A and B) or from ZW to XY (neo-Y frequency shown by black lines, panels C and D) occur similarly regardless of sex ratio biases present before (B versus D) or after (C versus A, dashed lines show male frequency). 
During invasion by a neo-ZW sex determination system (A and B), the ancestral Y fixes in both males and females (blue and red lines). 
Similarly, the ancestral W allele fixes in males and females (blue and red lines) during a ZW to XY transition. 
In this plot, there is no gametic competition ($t^\female=t^\male=0$) and meiotic drive occurs during male meiosis only ($\alpha^\female_{\Delta}=0$, $\alpha^\Hermaphrodite_{\Delta}=-1/5$). Therefore, sex ratio biases can only arise when the \textbf{A} locus is linked to an XY sex-determining locus.
In panels A and C, the neo-sex-determining locus is more closely linked to the \textbf{A} locus than the ancestral sex-detemining region ($r=1/2$, $R=1/20$) such that a neo-Y can caused biased sex ratios (panel C).
In panels B and D, the ancestral sex-determining locus is more closely linked to the \textbf{A} locus than the neo-sex-detemining locus ($r=1/20$, $R=1/2$). 
Therefore, an ancestral XY sex determination can have a biased zygotic sex ratio that becomes unbiased after an unlinked neo-W invades (B). 
However, in panel D, a unlinked neo-Y invades an ancestral ZW sex determination system in a similar manner but no biases to the zygotic sex ratio occur. 
With diploid selection alone, neo-sex-determining loci do not spread if they are less closely linked to the \textbf{A} locus than the ancestral sex-determining locus (see equation \eqref{eq:lambda_neoW} and Figure \ref{fig:Combination_Centimorgans}A). 
In this plot there are no sex differences in selection and an equilibrium is maintained because selection in diploids opposes meiotic drive, $s^\female =s^\male = 1/5$, $h^\female = h^\male = 7/10$.
\textcolor{red}{Aethetic adjustments: Could add titles to the columns/rows: neo-W for row 1, neo-Y for row 3, $r=0.5$, $R=0.05$ for column 1 and $r=0.05$, $R=0.5$ for column 2. Could adjust padding (too much whitespace where there is no axis label). It also seems could increase ratio of font size relative to plot size to make figure more compact. Matt - could you uncomment the line legends in the Mathematica file (function not included in my Mathematica version).}
}
\label{fig:Combination_Turnover}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

>>>>

Table \ref{tab:haplotype_growth} illustrates a number of key points about the invasion of neo-Y and neo-W mutations. 
For a neo-Y, invasion is driven by the fitness of male gametes and diploids, where the latter is weighted by the chance that a female egg will give rise to that diploid. 
For example, matings with $A$-bearing female gametes occur with probability $p_{X}^{\female} w_{A}^{\female}/\bar{w}_{H}^{\female}$. 
Since a neo-Y is always found in males, the allele frequencies at the neo-Y (\textbf{M}) locus only change in males. 
Therefore, invasion by a neo-Y allele does not involve any female diploid selection terms ($w_{g}^{\female}$). 
Similarly, invasion by a neo-W is driven by the fitness of female gametes and diploids and does not involve any direct selection in male diploids.
However, in the case of a neo-W, female diploids can result from matings with either an X-bearing or a Y-bearing sperm. 
In either case, the zygote will develop as a female due to the presence of a neo-W. 
For example, neo-W females will inherit an $A$ allele from a male gamete with probability $\bar{p}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$, where $\bar{p}^{\male}=p_{Y}^{\male}q+p_{X}^{\male}(1-q)$ is the frequency of the A allele among both X- and Y-bearing male gametes. 
By contrast, females that do not carry the neo-W only result from matings with X-bearing sperm. 
They will therefore inherit an $A$ from a male gamete with probability $(1-q)\bar{p}_{X}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$. 
If the \textbf{A} locus is initially linked to the ancestral sex-determining locus, \textbf{X}, the frequency of the $A$ allele among X- and Y-bearing sperm can differ (equation \ref{eq:freq_diffs}). 
Thus, eggs with and without a neo-W differ in the frequency of $A$ alleles they obtain from mating with male gametes.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Heterogametic transitions with/without sex ratio change
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{Combination_Turnover}\\
\caption{
Heterogametic transitions from XY to ZW sex determination (neo-W frequency shown by black lines, panels A and B) or from ZW to XY (neo-Y frequency shown by black lines, panels C and D) occur similarly regardless of sex ratio biases present before (B versus D) or after (C versus A, dashed lines show male frequency). 
During invasion by a neo-ZW sex determination system (A and B), the ancestral Y fixes in both males and females (blue and red lines). 
Similarly, the ancestral W allele fixes in males and females (blue and red lines) during a ZW to XY transition. 
In this plot, there is no gametic competition ($t^\female=t^\male=0$) and meiotic drive occurs during male meiosis only ($\alpha^\female_{\Delta}=0$, $\alpha^\Hermaphrodite_{\Delta}=-1/5$). Therefore, sex ratio biases can only arise when the \textbf{A} locus is linked to an XY sex-determining locus.
In panels A and C, the neo-sex-determining locus is more closely linked to the \textbf{A} locus than the ancestral sex-detemining region ($r=1/2$, $R=1/20$) such that a neo-Y can caused biased sex ratios (panel C).
In panels B and D, the ancestral sex-determining locus is more closely linked to the \textbf{A} locus than the neo-sex-detemining locus ($r=1/20$, $R=1/2$). 
Therefore, an ancestral XY sex determination can have a biased zygotic sex ratio that becomes unbiased after an unlinked neo-W invades (B). 
However, in panel D, a unlinked neo-Y invades an ancestral ZW sex determination system in a similar manner but no biases to the zygotic sex ratio occur. 
With diploid selection alone, neo-sex-determining loci do not spread if they are less closely linked to the \textbf{A} locus than the ancestral sex-determining locus (see equation \eqref{eq:lambda_neoW} and Figure \ref{fig:Combination_Centimorgans}A). 
In this plot there are no sex differences in selection and an equilibrium is maintained because selection in diploids opposes meiotic drive, $s^\female =s^\male = 1/5$, $h^\female = h^\male = 7/10$.
\textcolor{red}{Aesthetic adjustments: Could add titles to the columns/rows: neo-W for row 1, neo-Y for row 3, $r=0.5$, $R=0.05$ for column 1 and $r=0.05$, $R=0.5$ for column 2. Could adjust padding (too much whitespace where there is no axis label). It also seems could increase ratio of font size relative to plot size to make figure more compact. Matt - could you uncomment the line legends in the Mathematica file (function not included in my Mathematica version).}
}
\label{fig:Combination_Turnover}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
